\documentclass[twocolumn]{article}
\usepackage{cite}
\usepackage[strings]{underscore}
\author{Donnacha Oisín Kidney}
\title{Outline For FYP: A Solver for Commutative Rings in Agda}
\begin{document}
\maketitle

Dependently typed programming languages, such as
Agda\cite{norell_dependently_2008}, Idris\cite{brady_idris_2013}, and
Coq\cite{the_coq_development_team_2018_1219885}, extend the type systems of
programming languages to allow for proofs to be written as programs. They are of
interest to mathematicians and programmers alike: as they're based on a
constructivist interpretation of mathematics, the languages themselves are
viable avenues for a formal foundation of modern mathematics; as programming
languages, many of the features and concepts that have been developed have
proven useful in mainstream languages.

Mathematics must be rebuilt from the ground-up in each of these languages. The
goal of this project is to build a tool for one particular area in the language
Agda.

The tool is an automated solver for equalities in commutative rings. Without the
solver, proofs of basic equalities can be hundreds of lines long, and difficult
to understand or write. Automating the process isn't as simple as just checking
if two sides of an equation are the same, though: the solver has to be
\emph{evidence-providing}, meaning that it constructs a formal proof of the
equality based on the ring axioms.

Such a solver already exists for Agda\cite{danielsson_agda_2018}: the target
implementation for this project is a highly optimized version of the existing
solver, which dramatically improves the efficiency. The optimizations are the
same as those in Coq's ring solver\cite{gregoire_proving_2005}, although the
implementation strategy is different, and more properties are verified.
\bibliographystyle{IEEEtranS}
\bibliography{../horners-rule.bib}
\end{document}